---
title: "Reproducible Research Project1"
author: "Sachin"
date: "15/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
1. Load the data (i.e. read.csv())

```{r}
get.data <- function (){
        if (!file.exists("data")) {
        message("Creating Data folder in working directory")
        dir.create("data")
        }

        if(!file.exists("data/repdata-data-activity")) {
        fileURL <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
        download.file(fileURL, destfile = "./data/repdata-data-activity.zip")
        file.name <- "./data/repdata-data-activity.zip"
        unzip(file.name, exdir = "data")
        unlink(file.name)
}
else message("data already exists")
}
get.data()
activity <- read.csv("./data/activity.csv", colClasses = c("numeric", "Date", "numeric"))
dim(activity)
```
There are 17568 records in the activity

2. Process/transform the data (if necessary) into a format suitable for your analysis

## What is mean total number of steps taken per day
1. Calculate the total number of steps taken per day
```{r}

```
After NA removal we have only 15264 records

2. Make a histogram of the total number of steps taken each day
```{r}
library(ggplot2)
library(dplyr)
library(lubridate)
library(lattice)
activity$date <-as.Date(activity$date,"%d-%m-%Y")
activityAgg <- aggregate(steps ~ date, activity, sum, na.rm = TRUE)
hist(activityAgg$steps, main = "Total steps by day", xlab = "day", col = "red")
```

3. Calculate and report the mean and median total number of steps taken per day
```{r}
median(activityAgg$steps)
```
Median 10765
```{r}
mean(activityAgg$steps)
```
Mean 10766.19

## What is the average daily activity pattern?
1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
time_series <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)

plot(row.names(time_series), time_series, type = "l", xlab = "5-min interval", 
    ylab = "Average across all Days", main = "Average number of steps taken", 
    col = "red")

```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_interval <- which.max(time_series)
names(max_interval)
```

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset
```{r}
activity_NA <- sum(is.na(activity))
activity_NA
```

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
StepsAverage <- aggregate(steps ~ interval, data = activity, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(activity)) {
    obs <- activity[i, ]
    if (is.na(obs$steps)) {
        steps <- subset(StepsAverage, interval == obs$interval)$steps
    } else {
        steps <- obs$steps
    }
    fillNA <- c(fillNA, steps)
}

new_activity <- activity
new_activity$steps <- fillNA
StepsTotal2 <- aggregate(steps ~ date, data = new_activity, sum, na.rm = TRUE)

```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}
hist(StepsTotal2$steps, main = "Total steps by day", xlab = "day", col = "red")

mean(StepsTotal2$steps)
median(StepsTotal2$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day
2. Make a panel plot containing a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r}
day <- weekdays(activity$date)
daylevel <- vector()
for (i in 1:nrow(activity)) {
    if (day[i] == "Saturday") {
        daylevel[i] <- "Weekend"
    } else if (day[i] == "Sunday") {
        daylevel[i] <- "Weekend"
    } else {
        daylevel[i] <- "Weekday"
    }
}
activity$daylevel <- daylevel
activity$daylevel <- factor(activity$daylevel)

stepsByDay <- aggregate(steps ~ interval + daylevel, data = activity, mean)
names(stepsByDay) <- c("interval", "daylevel", "steps")
xyplot(steps ~ interval | daylevel, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```
